---
title: "Christopher Mara OR Prof Dev Assign 1"
output:
  word_document: default
  html_document: default
---

```{r}
library(lubridate)
library(tidyverse)
library(s20x)
library(ggplot2)
foodstuffs.tibble = read_csv("Data/FoodstuffsDemand2026.csv")
head(foodstuffs.tibble)
```

```{r}
foodstuffs_pivot = foodstuffs.tibble |>
  pivot_longer(cols = c(2:57), names_to = "Date", values_to = "Demand")
head(foodstuffs_pivot)
```
```{r}
foodstuffs_pivot = foodstuffs_pivot |>
  mutate(Chain = ifelse(startsWith(Supermarket, "Four"), "Four Square", ifelse(startsWith(Supermarket, "Pak"), "Pak 'n Save", "New World"))) |>
  mutate(Date = as.Date(Date, format = "%d/%m/%Y"))
head(foodstuffs_pivot)
```

```{r}
foodstuffs_pivot |>
  group_by(Date) |>
  summarise(MeanDemand = mean(Demand, na.rm = TRUE)) |>
  ggplot() +
  aes(x = Date, y = MeanDemand) +
  geom_line(colour = "steelblue") +
  labs(title = "Average Demand Over Time",
       y = "Average Demand")
```

The average Demand over time across the stores cycles periodically, dropping to 0 every week, and peaking around 6


```{r}
foodstuffs_pivot |>
  group_by(Chain, Supermarket) |>
  summarise(MeanDemand = mean(Demand, na.rm = TRUE)) |>
  ggplot() + 
  aes(x = Chain, y = MeanDemand, fill = Chain) +
  geom_boxplot() +
  labs(title = "Average Demand per Supermarket", y = "Mean Demand", x = "Supermarket Chain")
head(foodstuffs_pivot)
```

Pak'n Save has the highest center, followed by New World, followed by Four Square
Pak'N Save has slightly larger spread
all three are slightly right-skewed

```{r}
foodstuffs2 = foodstuffs_pivot |>
  mutate(Date = as.Date(Date, format = "%d/%m/%Y"))
head(foodstuffs2)
head(foodstuffs_pivot)
```

```{r}
#mean per day
foodstuffs_pivot |>
  mutate(Day = wday(Date, label = TRUE)) |>
  group_by(Day) |>
  summarise(MeanDemand = mean(Demand)) |>
  ggplot() +
  aes(x = Day, y = MeanDemand, fill = Day) +
  geom_col(fill = "steelblue")
```
```{r}
foodstuffs_pivot = foodstuffs_pivot|>
  mutate(Day = wday(Date, label = TRUE)) |>
  mutate(Demand = ifelse(Demand >= 14 & Demand%%10 == 0, Demand/10, Demand)) |>
  mutate(Demand = ifelse(Demand >= 14, NA, Demand)) |>
  mutate(Demand = ifelse(Demand <= 1, NA, Demand))
foodstuffs_pivot = foodstuffs_pivot[foodstuffs_pivot$Day != "Sun", ]
head(foodstuffs_pivot)
foodstuffs_pivot = foodstuffs_pivot[foodstuffs_pivot$Date != as.Date("2026-06-01"), ]
```
```{r}
#mean per day
foodstuffs_pivot |>
  mutate(Day = wday(Date, label = TRUE)) |>
  group_by(Day) |>
  summarise(MeanDemand = mean(Demand, na.rm=TRUE)) |>
  ggplot() +
  aes(x = Day, y = MeanDemand, fill = Day) +
  geom_col(fill = "steelblue")
```
Possible ways to continue may be to create means + standard error for each store for each day of the week
```{r}
foodstuffs_pivot |>
  group_by(Supermarket) |>
  summarise(MeanDemand = mean(Demand, na.rm=TRUE), StanDev = sd(Demand, na.rm=TRUE))
```
```{r}
head(foodstuffs_pivot)
foodstuffs_pivot |>
  ggplot() +
  aes(x = Date, y = Demand, color = Supermarket) +
  geom_line() +
  labs(title = "Average Demand Over Time",
       y = "Average Demand") 
```


